---
title: "Using Tidyverse"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Load require libraries
```{r packages}
#use suppressMessages function to disable messages upon loading a package
suppressMessages(library(tidyverse))
suppressMessages(library(palmerpenguins))
```


## Task 1

### Question a
We cannot use the `read_csv()` function specifically to read this data because it expects the data to be comma-separated. The data in `data.txt` is separated by a different delimiter (semicolon). `read_csv()` does not allow specifying a different delimiter, so we must use `read_delim()` instead.
 
```{r Q1a}
#read in the data file
data <- read_delim("data/data.txt", delim = ";")

#have all variables have type "dbl"
data <- data |>
  mutate(across(everything(), as.double))

#display the data
data
```

### Question b
```{r Q1b}
#read in the data file
data2 <- read_delim("data/data2.txt", delim = "6",
                    col_types = "fdc") #'f' for fct, 'd' for dbl, 'c' for chr

#display the data
data2
```

\newpage 

## Task 2

### Question a
```{r Q2a}
#read in the data file
trailblazer <- read_csv("data/trailblazer.csv")

#glimpse of the data to confirm it was read in correctly 
glimpse(trailblazer)
```

\newpage
### Question b
```{r Q2b}
# Pivot the data to long format
trailblazer_longer <- trailblazer |> 
  pivot_longer(cols = -Player, #use all columns except Player
               names_to = c("Game", "Location"), 
               names_sep = "_",
               values_to = "Points"
               )
  
#display the first 5 rows with slice()
trailblazer_longer |> 
  slice(1:5)
```

### Question c
```{r Q2c}
avg_trailblazer <- trailblazer_longer |> 
  pivot_wider(names_from = Location,
              values_from = Points
              ) |> #creates the 90 x 4 tibble 
  group_by(Player) |> #turns into 9 x 4 tibble with each row for the 9 players
  summarise(mean_home = mean(Home, na.rm = TRUE),
            mean_away = mean(Away, na.rm = TRUE)
            ) |>
  mutate(difference = mean_home - mean_away) |>
  arrange(desc(difference))

#display the data
avg_trailblazer
```

From the tibble above, player **Jusuf Nurkic** scored more on average at home through the first 10 games of the season than away.




















