---
title: "Using Tidyverse"
author: "Calista Harris"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

Load require libraries

```{r packages}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
library(palmerpenguins)
```

## Task 1

### Question a

We cannot use the `read_csv()` function specifically to read this data because it expects the data to be comma-separated. The data in `data.txt` is separated by a different delimiter (semicolon). `read_csv()` does not allow specifying a different delimiter, so we must use `read_delim()` instead.

```{r Q1a}
#read in the data file
data <- read_delim("data/data.txt", delim = ";")

#have all variables have type "dbl"
data <- data |>
  mutate(across(everything(), as.double))
  
#display the data
data
```

### Question b

```{r Q1b}
#read in the data file
data2 <- read_delim("data/data2.txt", delim = "6",
                    col_types = "fdc") #'f' for fct, 'd' for dbl, 'c' for chr

#display the data
data2
```

\newpage

## Task 2

### Question a

```{r Q2a}
#read in the data file
trailblazer <- read_csv("data/trailblazer.csv")

#glimpse of the data to confirm it was read in correctly 
glimpse(trailblazer)
```

\newpage

### Question b

```{r Q2b}
#pivot the data to long format
trailblazer_longer <- trailblazer |> 
  pivot_longer(cols = -Player, #use all columns except Player
               names_to = c("Game", "Location"), 
               names_sep = "_",
               values_to = "Points"
               )
  
#display the first 5 rows with slice()
trailblazer_longer |> 
  slice(1:5)
```

### Question c

```{r Q2c}
avg_trailblazer <- trailblazer_longer |> 
  pivot_wider(names_from = Location,
              values_from = Points
              ) |> #creates the 90 x 4 tibble 
  group_by(Player) |> #turns into 9 x 4 tibble with each row for the 9 players
  summarise(mean_home = mean(Home, na.rm = TRUE),
            mean_away = mean(Away, na.rm = TRUE)
            ) |>
  mutate(difference = mean_home - mean_away) |>
  arrange(desc(difference))

#display the data
avg_trailblazer
```

From the tibble above, player **Jusuf Nurkic** scored more on average at home through the first 10 games of the season than away.

\newpage

## Task 3

### Question a

-   \<NULL\>: means that there are no values for that particular cell in the data table where there is a missing or empty entry. This is common in hierarchical data when certain groups don’t have data for some variables 

-   \<dbl[52]\>: means that the cell contains a list-column with 52 numeric values, indicating repeated or nested data. This is an example of hierarchical data stored in a rectangular format 

-   \<list\>: indicates that the cell contains a list-column that could contain any type of object. From other lists, this is common in tibbles when data is too complex or nested to fit into a single vector 

### Question b

```{r Q3b}
penguins |> 
  count(species, island) |> 
  pivot_wider(names_from = island, values_from = n,
              values_fill = 0 #from tibble, missing combinations are value 0
              ) |>
  group_by(species)
```

\newpage

## Task 4
```{r Q4}
complete_bill_length <- penguins |> 
  mutate(bill_length_mm = if_else(
      is.na(bill_length_mm) & species == "Adelie", 26,
      if_else(is.na(bill_length_mm) & species == "Gentoo", 30, bill_length_mm)
  )) |> 
  arrange(bill_length_mm) |> 
  slice(1:10)

#display the data
complete_bill_length
```











